How kids code and how we know it:\\An exploratory study in the Scratch project repository

### 2016-04-19 Paper/paper.tex
Added more info on procedures complexity

\begin{table*}[ht]
	\centering
	\begin{tabular}{lrrrrrr}
		\hline
		&\textbf{mean}&\textbf{min}&\textbf{Q1}&\textbf{median}&\textbf{Q3}&\textbf{max}\\
		\hline
		Sprites with code per project&5.68&1&1&2&5&525\\
		Scripts per project&17.35&1&2&5&12&3038\\
		Maximum indentation per project&4.56&1&3&3&5&488\\
		Lines of code (LOC) per project&154.55&1&12&29&76&34622\\
		LOC in Stage&4.80&0&0&0&3&2613\\
		LOC in Sprites&115.57&0&10&26&68&34613\\
		LOC in Procedures&34.17&0&0&0&0&20552\\
		McCabe Cyclomatic Complexity per script&1.58&0&0&0&0&246\\
		\hline
	\end{tabular}
	\caption{Size-related statistics from the dataset of \nScriptPrograms  non-empty Scratch projects}
	\label{tbl-size}
\end{table*}

@@

The lines of code metric was further analyzed to understand code placement. As shown in Table \ref{tbl-size}, the majority of Scratch code---74.78\% out of \nLOC lines of code---is written within sprites. An additional 3.1\% of the total lines are found in the stage class. More interestingly, the remaining 22.11\% are lines within defined procedures, which are found in only 7.7\% (17,979) of the projects.

As a complexity metric we also used the McCabe cyclomatic complexity \cite{mcCabe76},  a quantitative measure of the number of linearly independent paths through a program's source code. This is calculated per script by counting the number of decision points in the script plus one. In Scratch, decision points can be the \texttt{if} and \texttt{if else} blocks. The results of the cyclomatic complexity metric per script are plotted in Figure \ref{fig:cyclomatic}. The majority (78.33\%) of \nscripts scripts contain no decision points, while 13.08\% have a cyclomatic complexity of 2, containing exactly one decision point. The complexity is higher, over 4, for 3.67\% of the scripts. The analysis also highlighted 209 scripts with cyclomatic complexity over 100 and up to 246.

\noindent
\fbox{
	\begin{minipage}{0.46\textwidth}
		\emph{RQ1: The majority of Scratch projects are small and simple; 75\% of the projects have up to 5 sprites, 12 scripts, 76 lines of code and no decision points. Most code is written in sprites. Few projects (7.7\%) use procedures, but they use them a lot. There exist surprisingly large and complex projects.}
	\end{minipage}}

\subsection{Programming abstractions and concepts}
loops	
logic operations	
variables	
lists	
user interativity	
functions - arguments
parallelism	- wait blocks, briadcast, receive

>>>>

\begin{table*}[ht]
	\centering
	\begin{tabular}{lrrrrrr}
		\hline
		&\textbf{mean}&\textbf{min}&\textbf{Q1}&\textbf{median}&\textbf{Q3}&\textbf{max}\\
		\hline
		Sprites with code per project&5.68&1&1&2&5&525\\
		Scripts per project&17.35&1&2&5&12&3038\\
		Maximum indentation per project&4.56&1&3&3&5&488\\
		Lines of code (LOC) per project&154.55&1&12&29&76&34622\\
		LOC in Stage&4.80&0&0&0&3&2613\\
		LOC in Sprites&115.57&0&10&26&68&34613\\
		LOC in Procedures&34.17&0&0&0&0&20552\\
		McCabe Cyclomatic Complexity (CC) per script&1.58&1&1&1&1&246\\
		McCabe CC per Procedure script&3.75&1&1&2&4&183\\
		\hline
	\end{tabular}
	\caption{Size-related statistics from the dataset of \nScriptPrograms  non-empty Scratch projects}
	\label{tbl-size}
\end{table*}

@@

The lines of code metric was further analyzed to understand code placement. As shown in Table \ref{tbl-size}, the majority of Scratch code---74.78\% out of \nLOC lines of code---is written within sprites. An additional 3.1\% of the total lines are found in the stage class. More interestingly, the remaining 22.11\% are lines within defined procedures, which are found in only 7.7\% (17,979) of the projects. Those projects that contain procedures use them a lot; almost half of their total lines of code (48.81\%) are lines within procedures.

As a complexity metric we also used the McCabe cyclomatic complexity \cite{mcCabe76},  a quantitative measure of the number of linearly independent paths through a program's source code. This is calculated per script by counting the number of decision points in the script plus one. In Scratch, decision points can be the \texttt{if} and \texttt{if else} blocks. The results of the cyclomatic complexity metric per script are plotted in Figure \ref{fig:cyclomatic}. The majority (78.33\%) of \nscripts scripts contain no decision points, while 13.08\% have a cyclomatic complexity of 2, containing exactly one decision point. The complexity is higher, over 4, for 3.67\% of the scripts. The analysis also highlighted 209 scripts with cyclomatic complexity over 100 and up to 246\footnote{The Scratch identifiers of those projects can be found in \todo{link}}. Cyclomatic complexity was greater (mean value of 3.32\%) in defined procedures, with 56.46\% of the procedures having at least one decision point.

\noindent
\fbox{
	\begin{minipage}{0.46\textwidth}
		\emph{RQ1: The majority of Scratch projects are small and simple; 75\% of the projects have up to 5 sprites, 12 scripts, 76 lines of code and no decision points. Most code is written in sprites. Few projects (7.7\%) use procedures, but they use them a lot and for more complex code. There exist surprisingly large and complex projects.}
	\end{minipage}}

\subsection{Programming abstractions and concepts}

functions - arguments
variables

loops	
logic operations		
lists	
user interativity	
parallelism	- wait blocks, briadcast, receive

Control abstraction is the abstraction of actions while data abstraction is that of data structures.

Control abstraction involves the use of subprograms and related concepts control flows
Data abstraction allows handling data bits in meaningful ways. For example, it is the basic motivation behind datatype.
